---
title: "Shwet_DA"
output: html_document
---

#Shwet Jain
#SU ID: 368595022
#NET ID: skjain
#Assignment 1 


###Reading the dataset into the "emp" variable, declaring it as a dataframe and studying the data
```{r}
emp <- read.csv("employee_attrition.csv",sep="," ,header=TRUE,stringsAsFactors = FALSE)
emp <- data.frame(emp)
str(emp)
```
### Deleting some rows which don't add value to the Apriori Algorithm.

```{r}
del <- c(9,10,27,22) 
emp <- emp[,-del]
```

###Checking for NA values 
```{r}
sum(is.na(emp))

```

##Removing the 9 records that contain NA values and storing the dataframe in the new variable named   "clean_emp"
```{r}
clean_emp <- na.omit(emp)
```

###Deleting some rows which don't add value to the Apriori Algorithm.
```{r}

del1 <- c(13,19,21,24,25,26,28,29,30,31)
clean_emp <- clean_emp[,-del1]
```

###To check NA value doesn't exists
```{r}
sum(is.na(clean_emp))
```

#Descriptive Statistics 
##To study each and every variable's distribution 

###Boxplot Distribution of DailyRate to convert into a categorical variable 

```{r}
boxplot(clean_emp$DailyRate)
```


### Summary of DistanceFromHome to get an idea of the distribution. 

```{r}
summary(clean_emp$DistanceFromHome)
```


### Scatter Plot of Age to get an idea of the distribution.   

```{r}
plot(clean_emp$Age)
```

### Histogram of Education of Employees 

```{r}
hist(clean_emp$Education)
```

### Histogram of Environment Satisfaction to study the variable. 

```{r}
hist(clean_emp$EnvironmentSatisfaction) 
```

### Histogram of JobInvolvement to convert into a categorical variable 

```{r}
hist(clean_emp$JobInvolvement)
```


### Plotting Job Satisfaction  

```{r}
hist(clean_emp$JobSatisfaction)
```

### Summary of MonthlyIncome  

```{r}
summary(clean_emp$MonthlyIncome)
```


### Boxplot Distribution of MonthlyRate to convert into a categorical variable 

```{r}
boxplot(clean_emp$MonthlyRate)
```


### Histogram Distribution of WorkLifeBalance variable to convert into a categorical variable 

```{r}
boxplot(clean_emp$WorkLifeBalance)
```

#Data Pre-processing
##Making sure that the data is good to go for Apriori algorithm.

### Converting Education to a categorical variable

```{r}
giveLevel<-function(x)
{
  BucketsClean <- replicate(length(x),"Doctor")
  BucketsClean[x==1] <- "Below College"
  BucketsClean[x==2] <- "College"
  BucketsClean[x==3] <- "Bachelor"
  BucketsClean[x==4] <- "Master"
  return (BucketsClean)
}
clean_emp$Education <- giveLevel(clean_emp$Education)
```

### Converting JobSatisfaction, RelationshipSatisfaction, JobInvolvement, Environment Satisfaction variable into categorical variables.

```{r}
giveLevel1<-function(x)
{
  BucketsClean <- replicate(length(x),"Low")
  BucketsClean[x==2] <- "Medium"
  BucketsClean[x==3] <- "High"
  BucketsClean[x==4] <- "Very High"
  return (BucketsClean)
}

clean_emp$JobSatisfaction <- giveLevel1(clean_emp$JobSatisfaction)
clean_emp$RelationshipSatisfaction <- giveLevel1(clean_emp$RelationshipSatisfaction)
clean_emp$JobInvolvement <- giveLevel1(clean_emp$JobInvolvement)
clean_emp$EnvironmentSatisfaction <- giveLevel1(clean_emp$EnvironmentSatisfaction)
```

### Converting Age into a categorical variable.

```{r}
giveLevel2<-function(x)
{
  BucketsClean <- replicate(length(x),"Seniors")
  BucketsClean[x>18 & x<38] <- "Young"
  BucketsClean[x>37 & x<54] <- "Adults"
 
  return (BucketsClean)
}
clean_emp$Age <- giveLevel2(clean_emp$Age)
```

### Converting Daily Rate into a categorical variable
```{r}
giveLevel3<-function(x)
{
  BucketsClean <- replicate(length(x),"Low")
  BucketsClean[x>=500 & x<=1100] <- "Medium"
  BucketsClean[x>1100] <- "High"
  return (BucketsClean)
}
clean_emp$DailyRate <- giveLevel3(clean_emp$DailyRate)
```

### Converting Distance From Home into a categorical variable
```{r}
giveLevel4<-function(x)
{
  BucketsClean <- replicate(length(x),"Near")
  BucketsClean[x>=12 & x<=18] <- "Average"
  BucketsClean[x>=18] <- "Far"
  return (BucketsClean)
}
clean_emp$DistanceFromHome <- giveLevel4(clean_emp$DistanceFromHome)
```

### Converting Performance Rating into a categorical variable

```{r}
giveLevel5<-function(x)
{
  BucketsClean <- replicate(length(x),"Outstanding")
  BucketsClean[x==1] <- "Low"
  BucketsClean[x==2] <- "Good"
  BucketsClean[x==3] <- "Excellent"
  return (BucketsClean)
}
clean_emp$PerformanceRating <- giveLevel5(clean_emp$PerformanceRating)
```

##Converting WorkLifeBalance into a categorical variable

```{r}
giveLevel6<-function(x)
{
  BucketsClean <- replicate(length(x),"Very High")
  BucketsClean[x==1] <- "Bad"
  BucketsClean[x==2] <- "Medium"
  BucketsClean[x==3] <- "High"
  return (BucketsClean)
}
clean_emp$WorkLifeBalance <- giveLevel6(clean_emp$WorkLifeBalance)
```

##Converting Hourly Rate variable into a categorical variable

```{r}
giveLevel7<-function(x)
{
  BucketsClean <- replicate(length(x),"High")
  BucketsClean[x<50] <- "Low"
  BucketsClean[x>50 & x<80] <- "Average"
  return (BucketsClean)
}
clean_emp$HourlyRate <- giveLevel7(clean_emp$HourlyRate)
```

###onverting Monthly Rate variable into a categorical variable

```{r}
giveLevel8<-function(x)
{
  BucketsClean <- replicate(length(x),"High")
  BucketsClean[x<10000] <- "Low"
  BucketsClean[x>10000 & x<20000] <- "Average"
  return (BucketsClean)
}
clean_emp$MonthlyRate <- giveLevel8(clean_emp$MonthlyRate)
```

###Converting Monthly Income to a categorical variable

```{r}
giveLevel9<-function(x)
{
  BucketsClean <- replicate(length(x),"High")
  BucketsClean[x<3000] <- "Low"
  BucketsClean[x>3000 & x<7000] <- "Average"
  return (BucketsClean)
}
clean_emp$MonthlyIncome <- giveLevel9(clean_emp$MonthlyIncome) 
```

##Converting all the varibale into Factors so as to run the Apriori model 
 
```{r}    
clean_emp$Age <- as.factor(clean_emp$Age)
clean_emp$Attrition <- as.factor(clean_emp$Attrition)
clean_emp$BusinessTravel <- as.factor(clean_emp$BusinessTravel)
clean_emp$DailyRate <- as.factor(clean_emp$DailyRate)
clean_emp$Department <- as.factor(clean_emp$Department)
clean_emp$DistanceFromHome <- as.factor(clean_emp$DistanceFromHome)
clean_emp$Education <- as.factor(clean_emp$Education)              
clean_emp$EducationField <- as.factor(clean_emp$EducationField)     
clean_emp$EnvironmentSatisfaction <- as.factor(clean_emp$EnvironmentSatisfaction)
clean_emp$Gender <- as.factor(clean_emp$Gender)        
clean_emp$HourlyRate <- as.factor(clean_emp$HourlyRate)           
clean_emp$JobInvolvement <- as.factor(clean_emp$JobInvolvement)     
clean_emp$JobRole <- as.factor(clean_emp$JobRole)           
clean_emp$JobSatisfaction <- as.factor(clean_emp$JobSatisfaction)       
clean_emp$MaritalStatus <- as.factor(clean_emp$MaritalStatus)   
clean_emp$MonthlyIncome <- as.factor(clean_emp$MonthlyIncome)         
clean_emp$MonthlyRate <- as.factor(clean_emp$MonthlyRate)         
clean_emp$OverTime <- as.factor(clean_emp$OverTime)
clean_emp$PerformanceRating <- as.factor(clean_emp$PerformanceRating)
clean_emp$RelationshipSatisfaction <- as.factor(clean_emp$RelationshipSatisfaction)
clean_emp$WorkLifeBalance <- as.factor(clean_emp$WorkLifeBalance)
```


### Checking whether all the variables are converted to factors
```{r}
str(clean_emp)
```

### Installing packages and libraries for apriori algorithm
```{r}

library(arules)


library(arulesViz)
```


### Running the Default algorithm 

```{r}
rules_1<-apriori(clean_emp)
rules_1 <- sort (rules_1, by="confidence", decreasing=TRUE)
inspect(head(rules_1,10))
plot(rules_1,jitter=0)
```


### Running the algorithm  with testing values for Support and Confidence and displaying the result in the descending order of confidence.

```{r}
rules_2 <- apriori(clean_emp, parameter = list(supp = 0.5, conf = 0.6, minlen=3))
rules_2 <- sort(rules_2, by="confidence", decreasing=TRUE)
inspect(head(rules_2,5))
```

###Plotting

```{r}
plot(rules_2,jitter=0)
```


##Output and present the most interesting and significant rules which could predict "Attrition"


### Running the algorithm with testing values for Support and Confidence and displaying the result in the descending order of confidence.
### At the same time, controlling the RHS i.e Attrition output : "Yes", "No"

```{r}
rules_3 <- apriori(clean_emp, parameter=list (supp=0.4,conf = 0.8, minlen=3), appearance = list (default="lhs",rhs=c("Attrition=Yes","Attrition=No")),control = list(verbose = F))
rules_3 <- sort (rules_3, by="confidence",decreasing=TRUE)
inspect(sort(rules_3, by = "lift", decreasing = T))

```

#Interesting rule no.1 
 
## {OverTime=No, WorkLifeBalance=High } => {Attrition=No}                           
##  Support: 0.4061697, Confidence: 0.9080460, Lift: 1.0769204, Count: 474

##This means that if the worker doesn't work overtime and his work-life balance is high, then the attrition accounts to NO


#Interesting rule no.2

## {DistanceFromHome=Near, OverTime=No}  => {Attrition=No}                            ##  Support:0.4661525, Confidence: 0.9036545, Lift = 1.0717122, Count= 544

##This means that if the worker doesn't work overtime and his work-life balance is high, then the attrition accounts to NO




#Plotting

```{r}
plot(rules_3,jitter=0)
```


#Shiny App
#LINK

#https://skjain.shinyapps.io/IST_707_ShwetJain/

##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
##xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

#End of Assignment 
